\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2014-09-04 doc/sw-system.tex
Finished discussion section, closed #44
Responses to discussion feedback from Anu and AV #68

We now quantify the performance of the \sw system in terms of the recovery of
the training set images. \Fref{fig:results:sample:roc} shows receiver
operating characteristic (ROC) curves for CFHTLS stage 1 and stage 2. These
plots show the true positive rate (TPR), the number of sims correctly detected
divided by the total number of sims in the training set, and the false
positive rate (FPR), the number of duds incorrectly detected  divided by total
number of duds in the training set, both for a given sample of detections defined by
a particular probability threshold, which varies along the curves.  In both
stages, these curves show that true positive rates of  around 90\% were
achieved, at very low false positive rates. For comparison we show the results
of an analysis where the classifications of  training images were ignored, and
none of the agents allowed to learn: they were instead assigned initial values
of their confusion matrices of 0.75 for each element, which then remained
constant. This emulates a very simple unweighted voting scheme, where all
classifications are treated equally. In this case, the TPR remains under 80\%
in stage 1 and 60\% in stage 2, thus quantifying 
the benefit of including training
images and allowing the agents to learn. The choice of initial confusion
matrix is not very important: the same 0.75 initial values applied to normal,
learning agents results in a slightly lower TPR than the default case,
indicating that initializing the system less conservatively results in
slightly worse performance.

The dot-dashed curves show the impact of the offline analysis. At stage 1 the
results are very similar to the online version that was actually run (solid
line). However, at stage 2 there seems to be somewhat greater benefit to doing
the analysis in this way: over 85\% TPR is achieved at zero FPR.
Alternatively, if one is willing to accept a false positive rate of 5\%, the
true positive rate rises to over 95\%, showing that some of the sims that were
missed in the online analysis can be recovered by doing the analysis offline.
The same is true at stage 1, but to a lower degree.

@@

\todo{Phil}{write discussion (\#44).}

@@

We investigated the source of the incompleteness and impurity visible in 
\Fref{fig:results:sample:CP}, by examining the stage 2 false negatives and
false positives. \Fref{fig:discuss:performance:trajectories} shows 2 example
simulated lenses  that were missed ($\pr(\LENS|C,T) < 10^{-5}$) by the \sw
system, and 2 example non-lenses that were incoreectly flagged as candidates
($\pr(\LENS|C,T) > 0.95$) by the \sw system.

In some cases the rejection of the false negatives is understandable: the
lensed features are faint, or otherwise difficult to see. However, in other
cases a reasonably obvious lens was passed over.  This mainly seems to be due
to noise in the system: when only low-skill classifiers view a subject, all
the updates to its posterior probability are small, and if none are very
confident about the presence of lensing, the subject follows a random walk
down its trajectory plot. This can be seen in the top left panel of the
figure.  The false positives show similar behaviour, for example in the bottom
lefthand panel of the figure. As well as subjects being ``unlucky'' in this
way, there are two less common failure modes associated with mistakes made by
higher skill users, illustrated in the righthand column of the figure. In the
false negative subject shown in the top right-hand panel, several high skill
classifiers update the subject upwards in log probability by some way each
time, but other, comparable skill classifiers mis-classify the system to lower
probability. If we had retired subjects after crossing a high probability
detection threshold, this subject would have been correctly identified as a
lens candidate; instead, it was kept in the system and eventually scattered
down to retirement. The trajectory looks like a random walk, but with bigger
steps. The bottom righthand panel shows an example of a final, apparently
rare, failure mode: we see some short-step random walk behaviour, followed by
a mis-classification by a very high skill ``expert'' classifier after 20
classifications that promotes the subject to high probability.

@@

Mistakes by both high and low skill classifiers could be reduced by improving
the training in the system, which could be done in several ways. One is to 
make more training images available to those who want or need it. A basic
level of training images are needed for SWAP to build up an accurate picture
of each classifier's skill -- but one could imagine volunteers {\it choosing}
to see more training images (still at random) in their stream. Another could
be to provide more information about what lenses do. In this project, the
Spotter's Guide was always available on the site, but as a passive background
resource. We might consider providing more links to this guide in the feedback
messages shown to the classifiers as they go -- or perhaps extending these
messages to themselves include more explanations and example images.

@@

However, most identification failures (around 60\%) seem to be due to the
random walk problem. While improved training will help reduce this effect, it
is here that targeted task assignment could help improve system performance.
We advertised stage 2 of the current project to all registered users; it was
taken up by stage 1 veterans with a broad range of skill, and also picked up a
significant number of new users who did not have enough time to aquire high
skill (since stage 2 was quite short): of the 1964 volunteers who took part in
the stage 2 classification round, only 774 were veterans from stage 1.
\Fref{fig:discuss:performance:stage1vstage2} shows how the stage 1 skill maps
on to stage 2 skill and contribution.

This figure suggests that there may be some benefit to opening stage 2 on the
basis of agent skill, in order to reduce the noise in the system generated by
new and low skill volunteers in this more difficult classification stage. A
simple implementation of this would be to switch the web app to the stage 2
version when either a certain stage 1 skill (and/or effort) threshold was
passed. Good strategy would probably be to switch back to stage 1 after some
number of stage 2 classifications (to avoid losing the high skill users from
stage 1 entirely), and also to make it very clear which stage the classifier
was at (since stage 2 requires more careful attention). This is something we
plan to experiment with in future.

Finally, there could be some performance gains to be made by improving the
SWAP agent model, or its implementation. In a future paper we plan to
investigate the use of the test images as well as the training images in
learning the agent confusion matrix elements (Davis et al, in preparation).
One might also consider looking at introducing more conditional dependences in
these elements, to allow for some classifiers having greater skill in one type
of lens than another, or, more generally, as a function of lens property (such
as color, brightness, and image separation). In the current model, all agents
are considered to be completely independent, whereas in fact we might expect
there to be significant clustering of the confusion matrix elements. A
hierarchical model for the crowd, with hyper-parameters describing the
distribution of confusion matrix elements across the population, may well
accelerate the agent learning process by including the notion of one agent
being likely to be similar to its neighbours. Finally, it is worth noting that
the model of \citet{IBCC} explicitly avoids the assumption of the agent
confusion matrix elements being constant in time, allowing the development of
volunteer skill to be more accurately tracked -- and that they did see some
time-evolution in the supernova zoo classifiers' skill. Finding a way to
incorporate such a learning model into SWAP, while retaining its statistically
online character is an interesting challenge for future work.

>>>>

We now quantify the performance of the \sw system in terms of the recovery of
the training set images. At stage 1, this set contains around 5712 simulated
lenses, and 450 duds; at stage 2, we used 152 simulated lenses and 201 duds.
\Fref{fig:results:sample:roc} shows receiver operating characteristic (ROC)
curves for CFHTLS stage 1 and stage 2. These plots show the true positive rate
(TPR), the number of sims correctly detected divided by the total number of
sims in the training set, and the false positive rate (FPR), the number of
duds incorrectly detected  divided by total number of duds in the training
set, both for a given sample of detections defined by a particular probability
threshold, which varies along the curves.  In both stages, these curves show
that true positive rates of  around 90\% were achieved, at very low false
positive rates. For comparison we show the results of an analysis where the
classifications of  training images were ignored, and none of the agents
allowed to learn: they were instead assigned initial values of their confusion
matrices of 0.75 for each element, which then remained constant. This emulates
a very simple unweighted voting scheme, where all classifications are treated
equally. In this case, the TPR remains under 80\% in stage 1 and 60\% in stage
2, thus quantifying  the benefit of including training images and allowing the
agents to learn. The choice of initial confusion matrix is not very important:
the same 0.75 initial values applied to normal, learning agents results in a
slightly lower TPR than the default case, indicating that initializing the
system less conservatively results in slightly worse performance.

The dot-dashed curves show the impact of the offline analysis. At stage 1 the
results are very similar to the online version that was actually run (solid
line). However, at stage 2 there is marginal evidence of there being greater
benefit to doing the analysis offline. Over 85\% TPR is achieved at zero FPR
in the offline analysis , while if one is willing to accept a false positive
rate of 5\%, the true positive rate rises to over 95\%, showing that some of
the sims that were missed in the online analysis may be being recovered by
doing the analysis offline. The same is true at stage 1, but to a lower
degree. Assuming Poisson statistics for the fluctuations in the numbers of
recovered lenses, the uncertainty in the measured stage 2 TPR values is around
8\%, but the online and offline samples are highly correlated, such that the
uncertainty on the difference between the ROC curves is somewhat less than
this. Still, a larger validation set is needed to test these algorithmic
choices more rigorously. At stage 1, high TPR can be measured to better than
1\%.

@@

@@

We investigated the source of the incompleteness and impurity visible in 
\Fref{fig:results:sample:CP}, by examining the stage 2 false negatives and
false positives, and their behaviour as they are classified using the online
analysis trajectory plots introduced in \Sref{sec:swap}
(\Fref{fig:swap:subject-trajectories}).
\Fref{fig:discuss:performance:trajectories} shows 2 example simulated lenses 
that were missed ($\pr(\LENS|C,T) < 10^{-5}$) by the \sw system, and 2 example
non-lenses that were incorrectly flagged as candidates ($\pr(\LENS|C,T) >
0.95$) by the \sw system. While the final stage 2 analysis was done offline,
these trajectory plots still serve to illustrate the approximate
classification histories of the subjects.

In some cases the rejection of the false negatives is understandable: the
lensed features are faint, or in some cases, somewhat unrealistic as lenses.
However, in other cases a reasonably obvious lens was passed over.  This
mainly seems to be due to noise in the system: when only low-skill classifiers
view a subject, all the updates to its posterior probability are small, and if
none are very confident about the presence of lensing, the subject follows a
random walk down its trajectory plot. This can be seen in the top left panel
of the figure.

The false positives show similar behaviour, for example in the bottom
lefthand panel of the figure. As well as subjects being ``unlucky'' in this
way, there are two less common failure modes associated with mistakes made by
higher skill users, illustrated in the righthand column of the figure. In the
false negative subject shown in the top right-hand panel, several high skill
classifiers update the subject upwards in log probability by some way each
time, but other, comparable skill classifiers mis-classify the system to lower
probability. If we had retired subjects after crossing a high probability
detection threshold, this subject would have been correctly identified as a
lens candidate; instead, it was kept in the system and eventually scattered
down to retirement. The trajectory looks like a random walk, but with bigger
steps. The bottom righthand panel shows an example of a final, apparently
rare, failure mode: we see some short-step random walk behaviour, followed by
a mis-classification by a very high skill ``expert'' classifier after 20
classifications that promotes the subject to high probability.

@@

Mistakes by both high and low skill classifiers could be reduced by improving
the training in the system, which could be done in several ways. One is to 
make more training images available to those who want or need it. A basic
level of training images are needed for SWAP to build up an accurate picture
of each classifier's skill -- but one could imagine volunteers {\it choosing}
to see more training images (still at random) in their stream. We could also
experiment with providing greater training rates early on for all volunteers;
the risk associated with this is that retention rates may drop if too few
``fresh'' test images are shown early on. Another way of improving the
training could be to provide more information about what lenses do. In this
project, the Spotter's Guide was always available on the site, but as a
passive background resource. We might consider providing more links to this
guide in the feedback messages shown to the classifiers as they go -- or
perhaps extending these messages to themselves include more explanations and
example images. We might also investigate a more dynamic Spotter's Guide: a
set of manipulable model lenses illustrating all the possible image
configurations that that those deflectors can make could help volunteers gain
understanding of what lensed features can look like very quickly. Such a toy
is under development.

@@

However, most identification failures (around 60\%) seem to be due to the
random walk problem. Short, random direction steps arise from low skill
classifiers, arising from classifier inexperience (\Sref{sec:results:crowd}). 
While improved training will help reduce this effect, it is here
that targeted task assignment could also help improve system performance, by
bringing higher skill volunteers in to the mix. We
advertised stage 2 of the current project to all registered users; it was
taken up by stage 1 veterans with a broad range of skill, and also picked up a
significant number of new users who did not have enough time to aquire high
skill (since stage 2 was quite short): of the 1964 volunteers who took part in
the stage 2 classification round, only 774 were veterans from stage 1.
\Fref{fig:discuss:performance:stage1vstage2} shows how the stage 1 skill maps
on to stage 2 skill and contribution.

This figure suggests that there may be some benefit to opening stage 2 on the
basis of stage 1 agent skill, in order to reduce the noise in the system
generated by new and low skill volunteers in this more difficult
classification stage.  A simple implementation of this would be to switch the
web app to the stage 2 version when either a certain stage 1 skill (and/or
effort) threshold was passed. Good strategy would probably be to switch back
to stage 1 after some number of stage 2 classifications (to avoid losing the
high skill users from stage 1 entirely), and also to make it very clear which
stage the classifier was at (since stage 2 requires more careful attention).
This is something we plan to experiment with in future.

Finally, there could be some performance gains to be made by improving the
SWAP agent model, or its implementation. Low skill is typically a result  of
inexperience (\Sref{sec:results:crowd}), but it could be the agent that is
inexperienced as well as the classifier or its agent. In a future paper we
plan to investigate the use of the test images as well as the training images
in accelerating the agent's learning (Davis et al, in preparation). We also
plan to investigate the use of offline analysis at Stage 1, for the same
reason (see also \Sref{sec:discuss:efficiency} below.) 
One might also consider looking at
introducing more conditional dependences in the confusion matrix elements, to
allow for some classifiers having greater skill in spotting one type of lens
than another, or, more generally, as a function of lens property (such as
color, brightness, and image separation). In the current model, all agents are
considered to be completely independent, whereas in fact we might expect there
to be significant clustering of the confusion matrix elements. A hierarchical
model for the crowd, with hyper-parameters describing the distribution of
confusion matrix elements across the population, may well accelerate the agent
learning process by including the notion of one agent being likely to be
similar to its neighbours. Finally, it is worth noting that the model of
\citet{IBCC} explicitly avoids the assumption of the agent confusion matrix
elements being constant in time, allowing the development of volunteer skill
to be more accurately tracked -- and that they did see some time-evolution in
the supernova zoo classifiers' skill. Finding a way to incorporate such a
learning model into SWAP, while retaining its statistically online character
is an interesting challenge for future work.
